Legion: Expressing Locality and Independence with Logical Regions

### 2012-07-01 doc/popl2013/langdef.tex

@@

\newcommand{\regionexpand}[0]{[r'_1/r_1, \ldots, r'_n/r_n]}
\newcommand{\rtriple}[3]{\left({#1},{#2},{#3}\right)}
\newcommand{\rsingle}[1]{\rtriple{#1}{\emptyset}{\top}}
\newcommand{\rtripsub}[3]{\rtriple{#1}{\Phi_{#2}}{C_{#3}}}
\newcommand{\envsub}[2]{\Gamma, \Phi_{#1}, C_{#2} \vdash}
\newcommand{\typeenv}[0]{\Gamma, \Phi, \Omega \vdash}

@@

\framebox{$\typeenv bv : bool$} \hspace{1cm}
\framebox{$\typeenv iv : int$} \\

\finfrule{\typeenv e : \langle T_1,T_2 \rangle}{\typeenv e\text{.1}\ : T_1} \hspace{1cm}
\finfrule{\typeenv e : \langle T_1,T_2 \rangle}{\typeenv e\text{.2}\ : T_2} \\

\finfrule{\Gamma(id) = T}{\typeenv id : T} \\

\framebox{$\typeenv \text{null }T@r : T@r$} \hspace{1cm}
\framebox{$\typeenv \text{new }T@r : T@r$} \\

\finfrule{\typeenv e : T@(\oton{r})}{\typeenv \text{isnull}(e) : bool} \\

\finfrule
{\begin{array}{l}
\typeenv e : T@(\otok{k}{r'}) \\
\forall i. \exists j, r'_i \leq r_j \in \Omega^* \\
\end{array}}
{\typeenv upregion(e,\oton{r}) : T@(\oton{r})} \\

\finfrule
{\begin{array}{l}
\typeenv e : T@(\otok{k}{r'}) \\
\forall j. \exists i, r_j \leq r'_i \in \Omega^* \\
\end{array}}
{\typeenv downregion(e,\oton{r}) : T@(\oton{r})} \\

\finfrule
{\begin{array}{l}
\typeenv e_1 : T@(\oton{r}) \\
\forall i, reads(r_i) \in \Phi^*\end{array}}
{\typeenv \text{read}(e_1) : T} \\

\finfrule
{\begin{array}{l}
\typeenv e_1 : T@(\oton{r}) \\
\typeenv e_2 : T \\
\forall i, writes(r_i) \in \Phi^*
\end{array}}
{\typeenv \text{write}(e_1, e_2) : bool} \\

\finfrule
{\begin{array}{l}
\Gamma(id) = (T_1, T_2), \emptyset \rightarrow T_1 \\
\typeenv e_1 : T_1@(\oton{r}) \\
\typeenv e_2 : T_2 \\
\forall i, reduces(id, r_i) \in \Phi^*
\end{array}}
{\typeenv \text{reduce}(id, e_1, e_2) : bool} \\

\framebox{$\typeenv \text{newcolor }r : \text{coloring}(r)$} \\

\finfrule{\begin{array}{l}
\typeenv e_1 : \text{coloring}(r) \\
\typeenv e_2 : T@r \\
\typeenv e_3 : int
\end{array}}
{\typeenv \text{color}(e_1, e_2, e_3) : \text{coloring}(r)} \\

\finfrule{\begin{array}{l}\typeenv e_1 : int \\ \typeenv e_2 : int\end{array}}{\typeenv e_1 + e_2 : int} \hspace{1 cm}
\finfrule{\begin{array}{l}\typeenv e_1 : int \\ \typeenv e_2 : int\end{array}}{\typeenv e_1 < e_2 : bool} \\

\finfrule{\begin{array}{l}\typeenv e_1 : T_1 \\ \Gamma' = \Gamma[id/T_1] \\ \Gamma', \Phi, \Omega \vdash e_2 : T_2 \end{array}}{\typeenv : \text{let}\ id : T_1 \text{in}\ e_2 : T_2} \\

\finfrule{\begin{array}{l}\typeenv e_1 : bool \\ \typeenv e_2 : T \\ \typeenv e_3 : T\end{array}}{\typeenv \text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3 : T} \\

\finfrule{
\begin{array}{l}
\Gamma(id) = \forall \otok{k}{r'}.(\oton{T}),\Phi_i \rightarrow T_r \\
\typeenv e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\
\Phi_i[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*
\end{array}}
{\typeenv id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]} \\

\finfrule{
\begin{array}{l}
\typeenv e_1 : \text{coloring}(r_p) \\
\Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{i \not= j} r_i * r_j \\
\Gamma, \Phi, \Omega' \vdash e_2 : T \\
r_i \not\in \textit{regions\_of}(\Gamma, T)
\end{array}}
{\typeenv \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T} \\

\finfrule{
\begin{array}{l}
T_1 = \exists \otok{k}{r'}. T_2\text{ where }\Omega_1 \\
\Omega \models \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\typeenv e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]
\end{array}}
{\typeenv \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1} \\

\finfrule{
\begin{array}{l}
T_1 = \exists \otok{k}{r'}. T_2\text{ where }\Omega_1 \\
\typeenv e_1 : T_1 \\
\Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \\
\Omega' = \Omega \wedge \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\Gamma', \Phi, \Omega' \vdash e_2 : T_3 \\
r_i \not\in \textit{regions\_of}(\Gamma, T_1, T_3)
\end{array}}
{\typeenv \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3} \\

@@

\newcommand{\opsenv}{M, L, H, S, C \vdash}
\newcommand{\opsenvm}[1]{{#1}, L, H, S, C \vdash}
\newcommand{\opsenvs}[1]{M, L, H, {#1}, C \vdash}
\newcommand{\opsenvls}[2]{M, {#1}, H, {#2}, C \vdash}
\newcommand{\opsenvms}[2]{{#1}, L, H, {#2}, C \vdash}
\newcommand{\opsenvmls}[3]{{#1}, {#2}, H, {#3}, C \vdash}

\framebox{$\opsenv e \mapsto v, S', E$}

@@

\framebox{$\opsenv bv \mapsto bv, []$} \hspace{1cm}
\framebox{$\opsenv iv \mapsto iv, []$} \\

\finfrule
{\opsenv e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenv e\text{.1} \mapsto v_1, E} \hspace{1cm}
\finfrule
{\opsenv e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenv e\text{.2} \mapsto v_2, E} \\

\finfrule
{L(id) = v}
{\opsenv id \mapsto v, []} \\

\framebox{$\opsenv \text{null }T@r \mapsto null, []$} \\

\finfrule
{\begin{array}{l}
l \in M(r) \\
H(l) = T[M]
\end{array}}
{\opsenv \text{new }T@r \mapsto l, []} \\

\finfrule
{\opsenv e \mapsto v, E}
{\opsenv \text{isnull}(e) \mapsto false, E} \\

\finfrule
{\opsenv e \mapsto null, E}
{\opsenv \text{isnull}(e) \mapsto true, E} \\

\finfrule
{\opsenv e \mapsto v, E}
{\opsenv \text{upregion}(e, \oton{r}) \mapsto v, E}

\finfrule
{\begin{array}{l}
\opsenv e \mapsto v, E \\
v' = \begin{cases}
v, & \text{if $v \not= null$ and $\exists i, v \in M(r_i)$}. \\
null, & \text{otherwise}.
\end{cases}
\end{array}}
{\opsenv \text{downregion}(e, \oton{r}) \mapsto v', E} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto l, E \\
S' = \text{apply}(S, E, C) \\
v = S'[l]
\end{array}}
{\opsenv \text{read}(e_1) \mapsto v, E \dplus [ read(l, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvs{S'} e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenv \text{write}(e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ write(l, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvs{S'} e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenv \text{reduce}(id, e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ reduce_{id}(l, v) ]} \\

\framebox{$\opsenv \text{newcolor }r \mapsto \emptyset, []$} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto K, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvs{S'} e_2 \mapsto l, E_2 \\
S'' = \text{apply}(S', E_2, C) \\
\opsenvs{S''} e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \}
\end{array}}
{\opsenv \text{color}(e_1, e_2, e_3) \mapsto K', E_1 \dplus E_2 \dplus E_3} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvs{S'} e_2 \mapsto v_2, E_2 \\
v' = v_1 + v_2
\end{array}}
{\opsenv e_1 + e_2 \mapsto v', E_1 \dplus E_2} \hspace{1cm}

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvs{S'} e_2 \mapsto v_2, E_2 \\
v' = \begin{cases}
true, & \text{if $v_1 < v_2$}. \\
false, & \text{otherwise}.
\end{cases}
\end{array}}
{\opsenv e_1 < e_2 \mapsto v', E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto v_1, E_1 \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvls{L'}{S'} e_2 \mapsto v_2, E_2
\end{array}}
{\opsenv \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto true, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvls{L'}{S'} e_2 \mapsto v_2, E_2
\end{array}}
{\opsenv \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto false, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvls{L'}{S'} e_3 \mapsto v_3, E_3
\end{array}}
{\opsenv \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E_1 \dplus E_3} \\

\finfrule
{\begin{array}{l}
\opsenv e_1 \mapsto v_1, E_1 \\
S_1 = \text{apply}(S, E_1, C) \\
\opsenv e_2 \mapsto v_2, E_2 \\
\ldots \\
S_n = \text{apply}(S_{n-1}, E_n, C) \\
task id[\otok{k}{r'}](f_1 : T_1, \ldots, f_n : T_n) : T_r, E = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \\
L' = \{ (f_1, v_1), \ldots, (f_n, v_n) \} \\
\opsenvmls{M'}{L'}{S_n} e_{n+1} \mapsto v_{n+1}, E_{n+1}
\end{array}}
{\opsenv id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E_1 \dplus \ldots \dplus E_{n+1}} \\

\finfrule{
\begin{array}{l}
\opsenv e_1 \mapsto K, E_1 \\
\rho_i = \{ l : (l, i) \in K \} \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvms{M'}{S'} e_2 \mapsto v, E_2
\end{array}}
{\opsenv \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E_1 \dplus E_2} \\

\finfrule{
\begin{array}{l}
\opsenv e_1 \mapsto v, E \\
\rho_i = M[r_i] \\
v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle
\end{array}}
{\opsenv \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] \mapsto v', E} \\

\finfrule{
\begin{array}{l}
\opsenv e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvmls{M'}{L'}{S'} e_2 \mapsto v_2, E_2
\end{array}}
{\opsenv \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E_1 \dplus E_2 } \\

@@

If
\begin{enumerate}
\item $\typeenv e : T$
\item $\opsenv e \mapsto v, E$
\end{enumerate}
we would like to prove that all of the dynamic effects described by $E$ are
included in the static privileges $\Phi$.  A judgement $E : \Phi$ means that
for all effects $\epsilon$ in $E$,
\begin{enumerate}
\item if $\epsilon = read(l, v)$, then for some $r$, $reads(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = write(l, v)$, then for some $r$, $writes(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = reduce(id, l, v)$, then for some $r$, $reduces(id,r) \in \Phi \wedge l \in M(r)$
\end{enumerate}

@@

\finfrule
{\begin{array}{l}
\typeenv e : T \\
\opsenv e \mapsto v, E \\
M \sim \Omega \hspace{1cm}
L \sim_H \Gamma \hspace{1cm}
S \sim H
\end{array}}
{
v \sim_H M \llbracket T \rrbracket \hspace{1cm}
E \sim H \hspace{1cm}
E : \Phi
}

@@

\finfrule
{\begin{array}{l}
\Gamma, \Phi_1, \Omega \vdash e_1 : T_1 \\
\Gamma, \Phi_2, \Omega \vdash e_2 : T_2 \\
\opsenv e_1 \mapsto v_1, E_1 \\
\opsenv e_2 \mapsto v_2, E_2 \\
\Phi_1 \#_M \Phi_2
\end{array}}
{E_1 \# E_2}

>>>>

% fun latex tricks to make typeenv and opsenv more friendly
\makeatletter
\define@key{typeenv}{G}{\def\typeenv@G{#1}}
\define@key{typeenv}{P}{\def\typeenv@P{#1}}
\define@key{typeenv}{O}{\def\typeenv@O{#1}}
\newcommand{\typeenvx}[1][]{
{
% default values
\def\typeenv@G{\Gamma}
\def\typeenv@P{\Phi}
\def\typeenv@O{\Omega}
\setkeys{typeenv}{#1}
\typeenv@G, \typeenv@P, \typeenv@O \vdash \,
}}
\newcommand{\typeenv}[3][]{\typeenvx[#1] {#2} : {#3}}
\define@key{opsenv}{M}{\def\opsenv@M{#1}}
\define@key{opsenv}{L}{\def\opsenv@L{#1}}
\define@key{opsenv}{H}{\def\opsenv@H{#1}}
\define@key{opsenv}{S}{\def\opsenv@S{#1}}
\define@key{opsenv}{C}{\def\opsenv@C{#1}}
\newcommand{\opsenvx}[1][]{
{
% default values
\def\opsenv@M{M}
\def\opsenv@L{L}
\def\opsenv@H{H}
\def\opsenv@S{S}
\def\opsenv@C{C}
\setkeys{opsenv}{#1}
\opsenv@M, \opsenv@L, \opsenv@H, \opsenv@S, \opsenv@C \vdash \,
}}
\newcommand{\opsenv}[4][]{\opsenvx[#1] {#2} \mapsto {#3}, {#4}}
\makeatother

@@

@@

\framebox{$\typeenv{bv}{bool}$} \hspace{1cm}
\framebox{$\typeenv{iv}{int}$} \\

\finfrule
{\begin{array}{l}
\typeenvx e_1 : T_1 \\
\typeenvx e_2 : T_2
\end{array}}
{\typeenvx \langle e_1, e_2 \rangle : \langle T_1, T_2 \rangle} \\

\finfrule{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.1}\ : T_1} \hspace{1cm}
\finfrule{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.2}\ : T_2} \\

\finfrule{\Gamma(id) = T}{\typeenvx id : T} \\

\framebox{$\typeenvx \text{null }T@r : T@r$} \hspace{1cm}
\framebox{$\typeenvx \text{new }T@r : T@r$} \\

\finfrule{\typeenvx e : T@(\oton{r})}{\typeenvx \text{isnull}(e) : bool} \\

\finfrule
{\begin{array}{l}
\typeenvx e : T@(\otok{k}{r'}) \\
\forall i. \exists j, r'_i \leq r_j \in \Omega^* \\
\end{array}}
{\typeenvx upregion(e,\oton{r}) : T@(\oton{r})} \\

\finfrule
{\begin{array}{l}
\typeenvx e : T@(\otok{k}{r'}) \\
\forall j. \exists i, r_j \leq r'_i \in \Omega^* \\
\end{array}}
{\typeenvx downregion(e,\oton{r}) : T@(\oton{r})} \\

\finfrule
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\forall i, reads(r_i) \in \Phi^*\end{array}}
{\typeenvx \text{read}(e_1) : T} \\

\finfrule
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\typeenvx e_2 : T \\
\forall i, writes(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{write}(e_1, e_2) : bool} \\

\finfrule
{\begin{array}{l}
\Gamma(id) = (T_1, T_2), \emptyset \rightarrow T_1 \\
\typeenvx e_1 : T_1@(\oton{r}) \\
\typeenvx e_2 : T_2 \\
\forall i, reduces(id, r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{reduce}(id, e_1, e_2) : bool} \\

\framebox{$\typeenvx \text{newcolor }r : \text{coloring}(r)$} \\

\finfrule{\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r) \\
\typeenvx e_2 : T@r \\
\typeenvx e_3 : int
\end{array}}
{\typeenvx \text{color}(e_1, e_2, e_3) : \text{coloring}(r)} \\

\finfrule{\begin{array}{l}\typeenvx e_1 : int \\ \typeenvx e_2 : int\end{array}}{\typeenvx e_1 + e_2 : int} \hspace{1 cm}
\finfrule{\begin{array}{l}\typeenvx e_1 : int \\ \typeenvx e_2 : int\end{array}}{\typeenvx e_1 < e_2 : bool} \\

\finfrule{\begin{array}{l}
\typeenvx e_1 : T_1 \\
\Gamma' = \Gamma[id/T_1] \\
\typeenvx[G=\Gamma'] e_2 : T_2
\end{array}}
{\typeenvx : \text{let}\ id : T_1 \text{in}\ e_2 : T_2} \\

\finfrule{\begin{array}{l}\typeenvx e_1 : bool \\ \typeenvx e_2 : T \\ \typeenvx e_3 : T\end{array}}{\typeenvx \text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3 : T} \\

\finfrule{
\begin{array}{l}
\Gamma(id) = \forall \otok{k}{r'}.(\oton{T}),\Phi_i \rightarrow T_r \\
\typeenvx e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\
\Phi_i[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*
\end{array}}
{\typeenvx id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]} \\

\finfrule{
\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r_p) \\
\Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{i \not= j} r_i * r_j \\
\typeenvx[O=\Omega'] e_2 : T \\
r_i \not\in \textit{regions\_of}(\Gamma, T)
\end{array}}
{\typeenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T} \\

\finfrule{
\begin{array}{l}
T_1 = \exists \otok{k}{r'}. T_2\text{ where }\Omega_1 \\
\Omega \models \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]
\end{array}}
{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1} \\

\finfrule{
\begin{array}{l}
T_1 = \exists \otok{k}{r'}. T_2\text{ where }\Omega_1 \\
\typeenvx e_1 : T_1 \\
\Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \\
\Omega' = \Omega \wedge \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \\
r_i \not\in \textit{regions\_of}(\Gamma, T_1, T_3)
\end{array}}
{\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3} \\

@@

\framebox{$\opsenv{e}{v}{E}$}

@@

\framebox{$\opsenvx bv \mapsto bv, []$} \hspace{1cm}
\framebox{$\opsenvx iv \mapsto iv, []$} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = apply(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \langle e_1, e_2 \rangle \mapsto \langle v_1, v_2 \rangle, E_1 \dplus E_2} \\

\finfrule
{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenvx e\text{.1} \mapsto v_1, E} \hspace{1cm}
\finfrule
{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenvx e\text{.2} \mapsto v_2, E} \\

\finfrule
{L(id) = v}
{\opsenvx id \mapsto v, []} \\

\framebox{$\opsenvx \text{null }T@r \mapsto null, []$} \\

\finfrule
{\begin{array}{l}
l \in M(r) \\
H(l) = T[M]
\end{array}}
{\opsenvx \text{new }T@r \mapsto l, []} \\

\finfrule
{\opsenvx e \mapsto v, E}
{\opsenvx \text{isnull}(e) \mapsto false, E} \\

\finfrule
{\opsenvx e \mapsto null, E}
{\opsenvx \text{isnull}(e) \mapsto true, E} \\

\finfrule
{\opsenvx e \mapsto v, E}
{\opsenvx \text{upregion}(e, \oton{r}) \mapsto v, E}

\finfrule
{\begin{array}{l}
\opsenvx e \mapsto v, E \\
v' = \begin{cases}
v, & \text{if $v \not= null$ and $\exists i, v \in M(r_i)$}. \\
null, & \text{otherwise}.
\end{cases}
\end{array}}
{\opsenvx \text{downregion}(e, \oton{r}) \mapsto v', E} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E \\
S' = \text{apply}(S, E, C) \\
v = S'[l]
\end{array}}
{\opsenvx \text{read}(e_1) \mapsto v, E \dplus [ read(l, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenvx \text{write}(e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ write(l, v) ]} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
\end{array}}
{\opsenvx \text{reduce}(id, e_1, e_2) \mapsto true, E_1 \dplus E_2 \dplus [ reduce_{id}(l, v) ]} \\

\framebox{$\opsenvx \text{newcolor }r \mapsto \emptyset, []$} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto l, E_2 \\
S'' = \text{apply}(S', E_2, C) \\
\opsenvx[S=S''] e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \}
\end{array}}
{\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E_1 \dplus E_2 \dplus E_3} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = v_1 + v_2
\end{array}}
{\opsenvx e_1 + e_2 \mapsto v', E_1 \dplus E_2} \hspace{1cm}

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = \begin{cases}
true, & \text{if $v_1 < v_2$}. \\
false, & \text{otherwise}.
\end{cases}
\end{array}}
{\opsenvx e_1 < e_2 \mapsto v', E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto true, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E_1 \dplus E_2} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto false, E_1 \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[L=L',S=S'] e_3 \mapsto v_3, E_3
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E_1 \dplus E_3} \\

\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S_1 = \text{apply}(S, E_1, C) \\
\opsenvx e_2 \mapsto v_2, E_2 \\
\ldots \\
S_n = \text{apply}(S_{n-1}, E_n, C) \\
task id[\otok{k}{r'}](f_1 : T_1, \ldots, f_n : T_n) : T_r, E = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \\
L' = \{ (f_1, v_1), \ldots, (f_n, v_n) \} \\
\opsenvx[M=M',L=L',S=S_n] e_{n+1} \mapsto v_{n+1}, E_{n+1}
\end{array}}
{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E_1 \dplus \ldots \dplus E_{n+1}} \\

\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
\rho_i = \{ l : (l, i) \in K \} \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[M=M',S=S'] e_2 \mapsto v, E_2
\end{array}}
{\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E_1 \dplus E_2} \\

\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto v, E \\
\rho_i = M[r_i] \\
v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle
\end{array}}
{\opsenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] \mapsto v', E} \\

\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1, C) \\
\opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E_1 \dplus E_2 } \\

@@

If
\begin{enumerate}
\item $\typeenvx e : T$
\item $\opsenvx e \mapsto v, E$
\end{enumerate}
we would like to prove that all of the dynamic effects described by $E$ are
included in the static privileges $\Phi$.  A judgement $E : \Phi$ means that
for all effects $\epsilon$ in $E$,
\begin{enumerate}
\item if $\epsilon = read(l, v)$, then for some $r$, $reads(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = write(l, v)$, then for some $r$, $writes(r) \in \Phi \wedge l \in M(r)$
\item if $\epsilon = reduce(id, l, v)$, then for some $r$, $reduces(id,r) \in \Phi \wedge l \in M(r)$
\end{enumerate}

@@

\finfrule
{\begin{array}{l}
\typeenvx e : T \\
\opsenvx e \mapsto v, E \\
M \sim \Omega \hspace{1cm}
L \sim_H \Gamma \hspace{1cm}
S \sim H
\end{array}}
{
v \sim_H M \llbracket T \rrbracket \hspace{1cm}
E \sim H \hspace{1cm}
E : \Phi
}

@@

\finfrule
{\begin{array}{l}
\typeenvx[P=\Phi_1] e_1 : T_1 \\
\typeenvx[P=\Phi_2] e_2 : T_2 \\
\opsenvx e_1 \mapsto v_1, E_1 \\
\opsenvx e_2 \mapsto v_2, E_2 \\
\Phi_1 \#_M \Phi_2
\end{array}}
{E_1 \# E_2}
